---
title: ggplot
editor: visual
engine: knitr
filters:
  - webr
---

<html>
<script type="module">

//=== WebR setup ===

import { WebR } from 'https://webr.r-wasm.org/latest/webr.mjs'; 
globalThis.webR = new WebR(); 
await globalThis.webR.init(); 
globalThis.webRCodeShelter = await new globalThis.webR.Shelter();
await globalThis.webR.installPackages(['ggplot2'])

</script>
</html>

::: {.callout-warning}
Installing and loading ggplot2 on webR takes a little while.
The install is happening in the background.
Don't worry, once you've waited to load the package everything else will be quick.
:::

### ggplot

* Very popular plotting package
* Good plots quickly
* Declarative - describe what you want not how to build it
* Contrasts w/Imperative - how to build it step by step
  
### Basics

* Load the package and some data

```{webr}
library(ggplot2)
```

* To build a plot using `ggplot` we start with the `ggplot()` function

```{webr}
ggplot()
```

* `ggplot()` creates a base ggplot object that we can then add things to - like a blank canvas
* We can also add optional arguments for information to be shared across different components of the plot
* The two main arguments we typically use here are 
* `data` - which is the name of the data frame we are working with, so `acacia`
* `mapping` - which describes which columns of the data are used for different aspects of the plot
* We create a `mapping` by using the `aes` function, which stands for "aesthetic", and then linking columns to pieces of the plot
* We'll start with telling ggplot what value should be on the x and y axes
* Let's plot the relationship betwen the circumference of an acacia and its height

```{webr}
ggplot(data = mtcars, mapping = aes(x = mpg, y = hp))
```

* This still doesn't create a figure, it's just a blank canvas and some information on
  default values for data and mapping columns to pieces of the plot
* We can add data to the plot using layers
* We do this by adding a `+` after the the `ggplot` function and then adding something called a `geom`, which stands for `geometry`
* To make a scatter plot we use `geom_point`

```{webr}
ggplot(data = mtcars, mapping = aes(x = mpg, y = hp)) +
  geom_point()
```
  
* To change things about the layer we can pass additional arguments to the `geom`
* We can do things like change
  * the `size` of the points, we'll set it to `3`
  * the `color` of the points, we'll set it to `"blue"`
  * the transparency of the points, which is called `alpha`, we'll set it to 0.5

```{webr}
ggplot(data = mtcars, mapping = aes(x = mpg, y = hp)) +
  geom_point(size = 3, color = "blue", alpha = 0.5)
```

* Try changing these values to make the graph look like you want it to

* To add labels (like documentation for your graphs!) we use the `labs` function

```{webr}
ggplot(data = mtcars, mapping = aes(x = mpg, y = hp)) +
  geom_point(size = 3, color = "blue", alpha = 0.5) +
  labs(x = "Miles per Gallon", y = "Horse Power",
       title = "Car stuff")
```

### Grouping

* Group on a single graph
* Look at influence of experimental treatment

```{webr}
ggplot(data = mtcars, mapping = aes(x = mpg, y = hp, color = cyl)) +
  geom_point(size = 3, alpha = 0.5)
```

* Try changing the above code to color based on the `gear`

* We can also split each group into different subplots (known as facets)

```{webr}
ggplot(data = mtcars, mapping = aes(x = mpg, y = hp)) +
  geom_point(size = 3, alpha = 0.5) +
  facet_wrap(~cyl)
```

* Try changing this code to create a subplot for each value in `vs` with points of size 4

::: {.callout-note}

## Exercise

Make a scatter plot with `hp` on the x axis and `wt` on the y axis. Label the x axis "Horse Power" and the y axis “Weight”.
Make one subplot for each value in `gear`.

```{webr}
# Add you code here
```

Your result should look like the plot below
:::

::: {.callout-tip collapse="true"}
## Expected Result

```{r}
#| echo: false
library(ggplot2)
ggplot(mtcars, aes(x = hp, y = wt)) +
  geom_point() +
  labs(x = "Horse Power", y = "Weight") +
  facet_wrap(~gear)
```
:::

::: {.callout-tip collapse="true"}
## Solution Code

```{r}
#| echo: true
#| eval: false
library(ggplot2)
ggplot(mtcars, aes(x = hp, y = wt)) +
  geom_point() +
  labs(x = "Horse Power", y = "Weight") +
  facet_wrap(~gear)
```
:::